---
title: "UCONN Clean Dataset and DEI Summary Tables"
author: "Chilombo Consulting, Vanessa"
date: "5/21/2023"
output: word_document
---

**_CLearing the environment_**
```{r include=FALSE}

rm(list = ls())
```


**_Setting working directory)**
```{r}
setwd("~/Documents/R/Chilombo Consulting/UCONN")
```

**Installing packages Loading libraries_**
```{r}

#packages already installed, if running for the first time add below
remotes::install_github("ddsjoberg/gtsummary")
install.packages("writexl")
#run libraries
require(readxl)
library(devtools)
library(dplyr)
library(tidyverse)
library(tidyr)
library(stringr)
library(padr)
library(gtsummary)
library(dtplyr)
library(writexl)




```

**_Determining the size of the dataset before opening it_**
```{r}
file.info('DEIData.xlsx')$size
#about 1 MB; not a big data venture by any means
```

**_Reading data_**
```{r}

DEIData_raw = read_xlsx("~/Documents/R/Chilombo Consulting/UCONN/DEIData.xlsx")
```


**Dropping rows for exclusion**
```{r}
 DEIData_master<-DEIData_raw %>% 
 #preview/test survey
   filter(DistributionChannel!="preview") %>% 
 #description column
   filter(DistributionChannel!="Distribution Channel") %>% 
 #pre-pharmacy students
  filter(is.na(Q1.5) |Q1.5 !="Yes, I am an undergraduate student who has NOT started my 1st year in the PharmD and/or a graduate program.") %>% 
  #filter out missing responses to Q2 onwards
  filter(!is.na(Q2))
```



**_Variable cleaning for each question that needs re categorization**
```{r}
DEIData_master_2<-DEIData_master %>% 
#create Q17 Race/Ethnicity groupings (URM/non-URM/White/Prfer not to Say) from Q17 
  mutate(`Q17_Race_Ethnicity` =  ifelse(Q17 =="White"  , "White",
                                 ifelse(Q17 == "Mixed Race/Biracial"  | Q17== "Black or African American" | 
                                        Q17== "Hispanic and/or Latino/a/e/x"|
                                        Q17=="Hispanic and/or Latino/a/e/x,White" |
                                        Q17=='Black or African American,East Asian/East Asian American',"URM",
                                ifelse(Q17 == "Prefer Not to Say",
                                      "Prefer Not to Say",
                                  ifelse( Q17 == "East Asian/East Asian American"  | Q17== "Middle Eastern/Northern African/West Asian" |
                                         Q17== "Middle Eastern/Northern African/West Asian,White"|
                                           Q17=="East Asian/East Asian   American,Other" | Q17=="South Asian/South Asian American"
                                         | Q17=="South Asian/South Asian American,White" |Q17== "East Asian/East Asian American,Other" |
                                           Q17=="East Asian/East Asian American,South Asian/South Asian American" | Q17=="Middle Eastern/Northern African/West Asian,White,Mixed Race/Biracial"
                                         , "Non-URM", NA
))))) %>% 


#create Other identity groupings from Q18)
    dplyr::mutate(Q18_FGCS=ifelse(grepl("First Generation College Student",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_FGI=ifelse(grepl("First Generation Immigrant",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_None=ifelse(grepl("None of the Above",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_LGBTQIA2S=ifelse(grepl("LGBTQIA2S+",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_Disabilities=ifelse(grepl("Individual with Disabilities \\(Mental or Physical\\)",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_TransferStudent=ifelse(grepl("Transfer Student",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_DeafHOH=ifelse(grepl("Deaf or Hard of Hearing",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_Other=ifelse(grepl("Other",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_PreferNottoSay=ifelse(grepl("Prefer Not to Say",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_SGI=ifelse(grepl("Second Generation Immigrant",Q18, ignore.case = T), "Yes","No")) %>% 
    dplyr::mutate(Q18_IS=ifelse(grepl("International Student",Q18, ignore.case = T), "Yes","No"))


     ## Check cleaning of race/ethnicity & other categories in console
      raceethnicituidentity_DQcheckDEIData_master_2 %>% count(Q17,Q17_Race_Ethnicity)
      otheridentity_DQcheck<-DEIData_master_2 %>% count(Q18,Q18_FGCS,Q18_FGI, Q18_None, Q18_LGBTQIA2S, Q18_Disabilities,Q18_TransferStudent,Q18_DeafHOH,Q18_Other,  Q18_PreferNottoSay,Q18_SGI, Q18_IS)


```

**_Variable renaming for questions of interest_**
```{r}
 DEIData_master_clean<-DEIData_master_2 %>% 
  # all unique response questions we need to use in the bivariate analysis
   dplyr::rename("Q1_Participant Type" = `Q1`,
                  "Q2_UCONN Addresses Issues Well" = `Q2`,
                 "Q4_SOP Addresses Issues Well" = `Q4`,
                 "Q8_Experienced_Stereotyping"= `Q8_1`,
               "Q8_Experienced_Microassaults"= `Q8_2`,
                "Q8_Experienced_Microinsults"=`Q8_3`,
                "Q8_Experienced_Microinvalidations"=`Q8_4`,
               "Q8_Experienced_Code-Switching"= `Q8_5`,
               "Q8_Experienced_Power Imbalance"=`Q8_6`,
                "Q8_Experienced_Other"=`Q8_7`,
                "Q7_Witnessed__Stereotyping"= `Q7_1`,
               "Q7_Witnessed_Microassaults"= `Q7_2`,
                "Q7_Witnessed__Microinsults"=`Q7_3`,
                "Q7_Witnessed_Microinvalidations"=`Q7_4`,
               "Q7_Witnessed_Code-Switching"= `Q7_5`,
               "Q7_Witnessed_Power Imbalance"=`Q7_6`,
                "Q7_Witnessed_Other"=`Q7_7`,
                "Q10_Material_Disrespectful"=`Q10 - Students`,
                "Q10_Material_Disrespectful_FS"=`Q10 - Faculty/Staff`,
                 "Q11_QA_Disrespectful"=`Q11 - Students`,
               "Q11_QA_Disrespectful_FS"= `Q11 - Faculty/Staff`,
               "Q14_Free_Expression"= `Q14`,
               "Q15_Heard_Expression"= `Q15`,
                 "Q16_Gender Identity"= `Q16`,
                "Q19_Identities_Represented_StudentBody"= `Q19`,
                "Q21_Identities_Represented_FacStaff"= `Q21`,
               "Q23_Representation_Important"= `Q23`,
               "Q24_Identities_Valued"= `Q24`,
               "Q25_Excel_Regardless_Identity"=`Q25`) %>% 
  #drop variables where no summary analysis is needed
    dplyr::select( -c('StartDate':'UserLanguage', 'Q1.5', 'Q3', 'Q5', 'Q6_8_TEXT','Q7_7_TEXT', 'Q8_7_TEXT', 'Q8_7_TEXT',
                      'Q9_10_TEXT', 'Q12_7_TEXT','Q13_28_TEXT','Q20', 'Q22', 'Q26_9_TEXT', 'Q28':'Q31')) %>% 

  #drop variables where select all that apply
    dplyr::select( -c('Q6','Q9', 'Q12', 'Q13', 'Q26')) %>% 

  #drop variables that were recategorized to other variable names for race/ethnicity and other identities
    dplyr::select(-c('Q17':'Q17_6_TEXT', 'Q18', 'Q18_6_TEXT')) 


```

**_Master Univariate Table_**
```{r}
 Univariate<-DEIData_master_clean 

 Univariate%>%
  tbl_summary(digits = ~ 0,
   missing_text = "(Missing)",
     sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**Univariate Table**") %>%  
  modify_caption("**Univariate Table*") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  #modify_footnote(all_stat_cols() ~ "missing values for Participant Type dropped; n (%)") %>% 
  bold_labels() %>% 
    as_gt() 
```

**_Master Participant_Type Table_**
```{r}
 Participant_Type<-DEIData_master_clean %>% 
    dplyr::select( -c('Q18_FGCS':'Q18_IS')) #remove all other identity groups from analysis

Participant_Type%>%
tbl_summary(by = `Q1_Participant Type`, digits = ~ 0,
   missing_text = "(Missing)", sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**Participant Type**") %>%  
  modify_caption("**Table 1. Participant Type**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  #modify_footnote(all_stat_cols() ~ "missing values for Participant Type dropped; n (%)") %>% 
  bold_labels() %>% 
    as_gt() 
```
**_Master Gender Table_**
```{r}
 Gender_Identity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q18_FGCS':'Q18_IS')) #remove all other identity groups from analysis

 Gender_Identity%>%
  tbl_summary(by = `Q16_Gender Identity`, digits = ~ 0,
   missing_text = "(Missing)", sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**Gender Identity**") %>%  
  modify_caption("**Table 2. Gender Identity**") %>%
     modify_footnote(everything() ~ NA) %>% #removes superscript from header
 # modify_footnote(all_stat_cols() ~ "missing values for Gender Identity dropped; n (%)") %>% 
  bold_labels() %>% 
    as_gt() 
```

**_Master Race and Ethnicity Table_**
```{r}
 Race_Ethnicity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q18_FGCS':'Q18_IS')) #remove all other identity groups from analysis
 Race_Ethnicity%>%
  tbl_summary(by = `Q17_Race_Ethnicity`, digits = ~ 0,
   missing_text = "(Missing)", sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**Race and Ethnicity Groups**") %>%  
  modify_caption("**Table 3. Race and Ethnicity Groups**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
 # modify_footnote(all_stat_cols() ~ "missing values for Race/Ethnicity Group dropped; n (%)") %>% 
  bold_labels() %>% 
    as_gt() 
```



**_Master Other Identity Tables for top 6 tables_**
```{r}
  FGCS_Identity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q18_FGI':'Q18_IS')) #remove other other identity groups from analysis

 FGCS_Identity%>%
  tbl_summary(by = `Q18_FGCS`, digits = ~ 0,
   missing_text = "(Missing)", sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**First Generation College Student Identity**") %>%  
  modify_caption("**Table 4. *First Generation College Student Identity**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  bold_labels() %>% 
    as_gt() 
 
  FGI_Identity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q18_FGCS', 'Q18_None':'Q18_IS')) #remove other other identity groups from analysis

 FGI_Identity%>%
  tbl_summary(by = `Q18_FGI`, digits = ~ 0,
   missing_text = "(Missing)",sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**First Generation Immigrant Identity**") %>%  
  modify_caption("**Table 5. *First Generation Immigrant Identity**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  bold_labels() %>% 
    as_gt() 
 
  
  LGBTQIA2S_Identity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q1_Participant Type','Q18_FGCS':'Q18_None','Q18_Disabilities':'Q18_IS')) #remove other other identity groups from analysis

LGBTQIA2S_Identity%>%
  tbl_summary(by = `Q18_LGBTQIA2S`, digits = ~ 0,
   missing_text = "(Missing)", sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**LGBTQIA2S+ Identity**") %>%  
  modify_caption("**Table 6. *LGBTQIA2S+ Identity**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  bold_labels() %>% 
    as_gt() 


  Disabilities_Identity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q1_Participant Type','Q18_FGCS':'Q18_LGBTQIA2S','Q18_TransferStudent':'Q18_IS')) #remove other other identity groups from analysis

  Disabilities_Identity%>%
  tbl_summary(by = `Q18_Disabilities`, digits = ~ 0,
   missing_text = "(Missing)",sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**Disabilities-mental or physical Identity**") %>%  
  modify_caption("**Table 7. Disabilities-mental or physical Identity**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  bold_labels() %>% 
    as_gt() 
  
 Transfer_Student_Identity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q1_Participant Type','Q18_FGCS':'Q18_Disabilities','Q18_DeafHOH':'Q18_IS')) #remove other other identity groups from analysis

 Transfer_Student_Identity%>%
  tbl_summary(by = `Q18_TransferStudent`, digits = ~ 0,
   missing_text = "(Missing)",sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**Transfer Student Identity**") %>%  
  modify_caption("**Table 8. Transfer Student Identity**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  bold_labels() %>% 
    as_gt() 
 
  SecondGenImm_Identity<-DEIData_master_clean %>% 
    dplyr::select( -c('Q1_Participant Type','Q18_FGCS':'Q18_PreferNottoSay','Q18_IS')) #remove other other identity groups from analysis

  SecondGenImm_Identity%>%
  tbl_summary(by = `Q18_SGI`, digits = ~ 0,
   missing_text = "(Missing)",sort = list(everything() ~ "frequency")) %>% 
  modify_header(label ~ "**Second Generation Immigrant Identity**") %>%  
  modify_caption("**Table 9.Second Generation Immigrant Identity**") %>%
  modify_footnote(everything() ~ NA) %>% #removes superscript from header
  bold_labels() %>% 
    as_gt() 
```

**_Final Clean Dataset_**
```{r}
write.table(DEIData_master_clean, paste ('Clean Dataset/UCONN_DEI_Clean_', Sys.Date(), ".txt"), sep ="\t", row.names=FALSE)

```

